Unlocking Accountability: A Mixed-Methods Analysis of Bottom-Up Approaches in Brazilian Public Policy

### 2023-10-27 Scripts/paper.Rmd
Bayesian fit + using gt to report results

[Insert Table]

@@

We estimate all models using the following regression equation:

$$
Y_i = \alpha + \beta_1 T_i + \gamma X_i + \theta Z_i + \epsilon_i
$$
where $i$ indexes the experiment units. $Y_i$ is $1$ if a construction work is finished, and $0$ otherwise, α is the intercept, β denotes the average treatment effect on the treated (ATT), and $T_i$ is a binary treatment indicator (municipality is part or not of Obra Transparent). $\gamma$ is a vector of fixed effects, $X_i$ is a matrix of Brazilian states’ fixed effects, $\theta$ is a vector of controls, and $Z_i$ an array of controls for the case $i$. The error term is denoted by $ε_i$. We cluster the standard errors at the municipality level as mayors are the main target of the local OSBs work. We fit a Bayesian model with rstanarm (Goodrich, 2023) with default weakly informative priors (Gelman et. al., 2008).

@@

```{r balancing process,  message=FALSE, echo=FALSE}

@@

df_balancing_aux %>%
  group_by(tipo_do_projeto) %>%
  summarise(n())

## pós projeto
df_balancing_aux_post <- obras_inicio_projeto_ot %>%
  mutate(municipio1 = tolower(iconv(municipio, "UTF-8",  "ASCII//TRANSLIT"))) %>%
  inner_join(aux_muni1, by = join_by(municipio1  == name_muni1, uf == abbrev_state)) %>%
  inner_join(ipea_final, by = join_by(code_muni == tcode)) %>%
  mutate(indicator_ot = ifelse(is.na(indicator_ot), 0, indicator_ot),
         perc_pea18 =pea18/populacao,
         log_pop = log(populacao),
         log_renda = log(rpc*populacao),
         log_rpc = log(rpc))
```
Table (ref number) summarizes the main results of the regression.

```{r balancing check, message=FALSE, echo=FALSE}
m.out <- matchit(indicator_ot ~ perc_pea18  + ext_pobres  + log_rpc + 
                  p10_ricos_desigualdade + freq_escola_15_1 + analfabetismo +
                  log_renda + log_pop + percentual_de_execucao + rede_de_ensino_publico + tipo_do_projeto , data = df_balancing,
                 replace = TRUE, ratio = 3)

summary(m.out)
# summary(m.out, un=FALSE)
# plot(m.out, type = "jitter", interactive = FALSE)
# 
# m.sum <- summary(m.out)
# 
# plot(m.sum, var.order = "unmatched")

# #eQQ plot
# plot(m.out, type = "qq", which.xs = ~ log_rpc + ext_pobres + p10_ricos_desigualdade)
# plot(m.out, type = "qq", which.xs = ~ perc_pea18 + log_pop + + freq_escola_15_1)
# 
# #eCDF plot - black line is treated, gray is control
# plot(m.out, type = "ecdf", which.xs = ~ log_rpc + ext_pobres + p10_ricos_desigualdade)
# plot(m.out, type = "ecdf", which.xs = ~  perc_pea18 + log_pop + + freq_escola_15_1)
# 
# #density plot - black line is treated, gray is control
# plot(m.out, type = "density", which.xs = ~ log_rpc + ext_pobres + p10_ricos_desigualdade)
# plot(m.out, type = "density", which.xs = ~  perc_pea18 + log_pop + + freq_escola_15_1)

# Full matching on a probit PS
m.out2 <- matchit(indicator_ot ~ perc_pea18  + ext_pobres  + log_rpc + 
                  p10_ricos_desigualdade + freq_escola_15_1 + analfabetismo +
                  log_renda + log_pop, data = df_balancing,
                 method = "full", distance = "glm", link = "probit")
# summary(m.out2, un = FALSE)
# m.sum2 <- summary(m.out2)
# plot(m.sum2, var.order = "unmatched")

## regression with matching

m.data <- match.data(m.out2) %>%
  dplyr::select(weights, code_muni, subclass ) %>%
  inner_join(df_balancing_aux, by = "code_muni") %>%
  mutate(indicator_ot = as.factor(indicator_ot))

fit <- lm(concluida ~ indicator_ot + perc_pea18  + ext_pobres  + log_rpc + 
                  p10_ricos_desigualdade + freq_escola_15_1 + analfabetismo +
                  log_renda + log_pop, data = m.data, weights = weights)

# avg_comparisons(fit,
#                 variables = "indicator_ot",
#                 vcov = ~ subclass,
#                 newdata = subset(m.data, indicator_ot == 1),
#                 wts = "weights") %>%
#   gt()
#

matching_data <- match.data(m.out2) %>%
  dplyr::select(weights, code_muni, subclass)

pre_dados <- df_balancing_aux %>%
  dplyr::select(id, concluida) %>%
  rename(concluida_pre = concluida) %>%
  mutate(id = as.character(id))

data_pos <- obras_fim_seg_fase %>%
  mutate(municipio1 = tolower(iconv(municipio, "UTF-8",  "ASCII//TRANSLIT"))) %>%
  inner_join(aux_muni1, by = join_by(municipio1  == name_muni1, uf == abbrev_state)) %>%
  inner_join(ipea_final, by = join_by(code_muni == tcode)) %>%
  mutate(indicator_ot = ifelse(is.na(indicator_ot), 0, indicator_ot),
         perc_pea18 =pea18/populacao,
         log_pop = log(populacao),
         log_renda = log(rpc*populacao),
         log_rpc = log(rpc)) %>%
  inner_join(matching_data, by = "code_muni") %>%
  inner_join(pre_dados, by = "id") %>%
  dplyr::filter(concluida_pre != concluida) %>%
  mutate(concluida = as.numeric(concluida))

fit1 <- lm(concluida ~ indicator_ot + perc_pea18  + ext_pobres  + log_rpc + 
                  p10_ricos_desigualdade + freq_escola_15_1 + analfabetismo +
                  log_renda + log_pop, data = data_pos, weights = weights)

@@

```{r tables 2, echo=FALSE,  message=FALSE, results='asis'}
avg_comparisons(fit1,
                variables = "indicator_ot",
                vcov = ~ subclass,
                newdata = subset(data_pos, indicator_ot == 1),
                wts = "weights") %>%
  gt()
```

@@

>>>>

@@

We estimate all models using the a varying-intercept regression equation (Gelman & Hill, 2006):

$$
Y_i = \alpha + \beta_1 T_i + \gamma X_i + \theta Z_i + \epsilon_i
$$
where $i$ indexes the construction units. $Y_i$ is $1$ if a construction work is finished, and $0$ otherwise, α is the intercept, β denotes the average treatment effect on the treated (ATT), and $T_i$ is a binary treatment indicator (municipality is part or not of Obra Transparent). $\gamma$ is a vector of varying-intercept, $X_i$ is a matrix of Brazilian municipalities random effects, $\theta$ is a vector of controls, and $Z_i$ an array of control variables for the case $i$. The error term is denoted by $ε_i$. We fit a Bayesian model with rstanarm (Goodrich, 2023) with default weakly informative priors on parameters (Gelman et. al., 2008).

@@

```{r prep data matching,  message=FALSE, echo=FALSE}

@@

df_balancing_aux1 <- df_balancing_aux %>%
  mutate(concluida_num = as.numeric(concluida))

df_balancing_aux_se <- df_balancing_aux1 %>%
  filter(uf.x %in% c("SP", "PR", "SP", "SC", "RS", "MG"))

pre_dados <- df_balancing_aux %>%
  dplyr::select(id, concluida) %>%
  rename(concluida_pre = concluida) %>%
  mutate(id = as.numeric(id))

data_pos <- obras_fim_seg_fase %>%
  mutate(municipio1 = tolower(iconv(municipio, "UTF-8",  "ASCII//TRANSLIT"))) %>%
  inner_join(aux_muni1, by = join_by(municipio1  == name_muni1, uf == abbrev_state)) %>%
  inner_join(ipea_final, by = join_by(code_muni == tcode)) %>%
  mutate(perc_pea18 =pea18/populacao,
         log_pop = log(populacao),
         log_renda = log(rpc*populacao),
         log_rpc = log(rpc)) %>%
  inner_join(pre_dados, by = "id") %>%
  dplyr::filter(concluida_pre == FALSE) %>%
  mutate(concluida_num = as.numeric(concluida))

df_balancing_aux_post_se <- data_pos %>%
  filter(uf.x %in% c("SP", "PR", "SP", "SC", "RS", "MG"))

@@

Table (ref number) summarizes the main results of the regression.

```{r pscore, message=FALSE, echo=FALSE}

# these are the no redundancy covariates
covs.nr <- c('perc_pea18', 'ext_pobres', 'log_rpc', 'p10_ricos_desigualdade', 'freq_escola_15_1', 'analfabetismo',
             'log_renda', 'log_pop', 'percentual_de_execucao', 'tipo_do_projeto')

# pscore estimation formula with original covariates

ps_spec <- formula(indicator_muni_ot ~ log_rpc  + ext_pobres + freq_escola_15_1 + analfabetismo +
                     log_renda   +  p10_ricos_desigualdade +
                       + percentual_de_execucao)

# Estimation of pscores using stan_glm
set.seed(1234)
ps_fit_pre <- stan_glm(ps_spec, family=binomial(link='logit'), 
                     data=df_balancing_aux_se, algorithm='optimizing', refresh=0, iter = 2000)

set.seed(1234)
ps_fit_pos <- stan_glm(ps_spec, family=binomial(link='logit'), 
                     data=df_balancing_aux_post_se, algorithm='optimizing', refresh=0, iter = 2000)
# extrqacting logit pscores
pscores_pre <- apply(posterior_linpred(ps_fit_pre), 2, mean)

pscores_pos <- apply(posterior_linpred(ps_fit_pos), 2, mean)

# matching
#matching without replacement, original formula

matches_pre <- matching(z=df_balancing_aux_se$indicator_muni_ot, score=pscores_pre, replace=FALSE)
matched_pre <- df_balancing_aux_se[matches$match.ind,]

matches_pos <- matching(z=df_balancing_aux_post_se$indicator_muni_ot, score=pscores_pos, replace=FALSE)
matched_pos <- df_balancing_aux_post_se[matches$match.ind,]

```{r fitting bayesian model, message=FALSE, echo=FALSE}

te_spec_nr <- update(ps_spec, concluida_num ~ indicator_muni_ot + (1 | municipio) + .)

#treatment effect without replacement

set.seed(20)
reg_ps_pre <- stan_lmer(te_spec_nr, data=df_balancing_aux_se[matches_pre$match.ind,])

## After
set.seed(20)
reg_ps_pos <- stan_lmer(te_spec_nr, data=df_balancing_aux_post_se[matches_pos$match.ind,])

summary(reg_ps_pre)['indicator_muni_ot', 1:5]
summary(reg_ps_pos)['indicator_muni_ot', 1:5]

```{r table 2, echo=FALSE,  message=FALSE, results='asis'}
# posterior draws

##
# Before
##

b1_pre_sims <- as.matrix(reg_ps_pre, 
                     pars = "indicator_muni_ot")

# Posterior mean and SD of treatment
b1_mean_pre <- mean(b1_pre_sims) # posterior mean
b1_sd_pre <- sd(b1_pre_sims)       # posterior SD

# Posterior median and 95% credible interval
b1_quant_pre <- quantile(b1_pre_sims,   probs = c(0.025, 0.50, 0.975))

b1_quant_pre <- data.frame(t(b1_quant_pre))
names(b1_quant_pre) <- c("Q2.5", "Q50", "Q97.5")

# Combine summary statistics of posterior simulation draws
b1_df_pre_cred <- data.frame(variables = "ATT before project",
                             post_mean = b1_mean_pre,
                             post_sd = b1_sd_pre,
                             b1_quant_pre[2], 
                             b1_quant_pre[1],
                             b1_quant_pre[3])

##
# After
##
b1_sims_after <- as.matrix(reg_ps_pos, 
                     pars = "indicator_muni_ot")

# Compute mean, SD, median, and 95% credible interval of treatment

# Posterior mean and SD of treatment
b1_mean_after <- mean(b1_sims_after) # posterior mean
b1_sd_after <- sd(b1_sims_after)       # posterior SD

# Posterior median and 95% credible interval
b1_quant_after <- quantile(b1_sims_after,   probs = c(0.025, 0.50, 0.975))

b1_quant_after <- data.frame(t(b1_quant_after))
names(b1_quant_after) <- c("Q2.5", "Q50", "Q97.5")

# Combine summary statistics of posterior simulation draws
b1_df_after_cred <- data.frame(variables = "ATT After project",
                             post_mean = b1_mean_after,
                             post_sd = b1_sd_after,
                             b1_quant_after[2], 
                             b1_quant_after[1],
                             b1_quant_after[3])
b1_df <- bind_rows(b1_df_pre_cred, b1_df_after_cred)

b1_df %>% gt() %>%
  tab_header(
    title = "Effect of local OSB social monitoring on construction delivery rates") %>%
  cols_label(
    variables = "Variables",
    post_mean = "Posterior mean",
    post_sd = "Posterior sd",
    Q50 = "Posterior median",
    Q2.5 = "Posterior 2.5% CI",
    Q97.5 = "Posterior 97.5% CI"
  ) %>%
  fmt_number(
    columns = c(post_mean, post_sd, Q50, Q2.5, Q97.5 ),
    decimals = 3,
    use_seps = FALSE
  ) %>%
  tab_source_note(
    source_note = "452 observations after matching, with municipalities of same states as participants of Obra Transparente."
  ) 
```

@@

Gelman, A., & Hill, J. (2006). Data analysis using regression and multilevel/hierarchical models. Cambridge university press.
